Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-29 discussion/discussion.tex
Updates from Overleaf

\chapter{Discussion}

XXXXXXXXXXX

@@

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks to learn low-dimensional music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

>>>>

\chapter{Conclusions}

In this work, we introduced a method that leverages self-supervised deep neural networks to learn low-dimensional music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information to analyze its performance in music segmentation tasks.

While not outperforming state-of-the-art results or traditional audio features in contrast to research such \cite{deepfeaturesegment, SalamonDeepSegmentation}, our musically-informed technique has significant potential for boundary detection tasks given its outstanding performance in accuracy metrics and overall performance improvement alongside the training set enlargement (GTZAN compared to Million Song Dataset). Most likely, so does for nearly all MIR downstream tasks that are not purely sonic-based.

The questions of whether such general-purpose audio representation can mimic human hearing \cite{Li2023MERT:Training, Turian2022HEAR:Representations} or whether these do decode underlying high-level musical content effectively remain scientifically unanswered given the lack of evaluation in that regard. That is a "thorn in our side" that we will pursue.

@@

### 2023-06-29 evaluation/evaluation.tex
Updates from Overleaf

@@

\item \textbf{Normalize the \textit{embeddiogram}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1 per dimension. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355embeddiogramnormalized.png}
    \caption[Embeddiogram. Track 355 (SALAMI dataset).]{Embeddiogram. Track 391 (SALAMI dataset).}
    \label{fig:image1}
  \end{minipage}

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 391 (SALAMI dataset).]{Novelty curve + peaks. Track 355 (SALAMI dataset).}
    \label{fig:image2}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix. Track 355 (SALAMI dataset).}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity lag matrix]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix. Track 355 (SALAMI dataset).}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix. Track 355 (SALAMI dataset).}
    \end{minipage}
\end{figure}

@@

Employing such embeddings in a traditional music segmentation algorithm can obtain state-of-the-art performance \cite{deepfeaturesegment}

@@

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of Music Structure Analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

All the evaluation experiments have been wrapped around MSAF \cite{NietoMSAF:FRAMEWORK}, a popular framework for Music Structure Analysis.

@@

Three primary metrics have been utilized to assess the model performance: Precision, Recall, and F-measure. Precision quantifies the proportion of accurately identified boundaries relative to all estimated boundaries to indicate the algorithm's accuracy in boundary detection.

>>>>

We have conducted experiments using a window duration of 4 seconds (\(w = 4 \times \text{{sr (Hz)}}\)). As a general rule of thumb, this duration corresponds to two 4/4 bars of a piece of music with a tempo where the quarter note equals 120 BPM.

@@

\item \textbf{Normalize the \textit{embeddiogram \ref{fig:embeddiogram}}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1 per dimension. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355embeddiogramnormalized.png}
    \caption[Embeddiogram. Track 355 (SALAMI dataset).]{Embeddiogram. Track 391 (SALAMI dataset).}
    \label{fig:embeddiogram}
  \end{minipage}

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 391 (SALAMI dataset).]{Novelty curve + peaks. Track 355 (SALAMI dataset).}
    \label{fig:novelty}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:SSM}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity lag matrix computation of the embeddiogram]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
        \label{fig:SSLM}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:Q}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:TBSM}
    \end{minipage}
\end{figure}

@@

Furthermore, these embeddiograms are a foundation for current state-of-the-art methods, which are expected to receive as input standard audio features. Consequently, they can be processed and manipulated like conventional features \ref{fig:SSM} \ref{fig:SSLM} \ref{fig:Q} \ref{fig:TBSM} \ref{fig:novelty}. Employing such embeddings in a traditional music segmentation algorithm can achieve state-of-the-art performance \cite{deepfeaturesegment}.

@@

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of music structure analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

All the evaluation experiments have been wrapped around MSAF \cite{MSAF}, a popular framework for MSA.

@@

Three primary metrics have been utilized to assess the model performance: Precision, Recall, and F-measure.

Precision quantifies the proportion of accurately identified boundaries relative to all estimated boundaries to indicate the algorithm's accuracy in boundary detection.

### 2023-06-29 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations, but one music piece can yield infinite interpretations, varying by factors such as instruments and performance styles.

The premise of our work is that regardless of the waveform or production style, abstract high-level features can be discerned, offering objective and thorough insight into a composition's musical content, analogous to analyzing sheet music for its basic behaviors.

@@

Despite differing cultural contexts, styles, time and key signatures, and sonic properties, we suggest a commonality between figures \ref{fig:mahler} and \ref{fig:giant_steps}. Both pieces display the same 'musicological fingerprint,' derived from their melodic and harmonic contour envelopes through non-diatonic major thirds.

@@

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/giant steps score.png}
        }
    \caption[Giant Steps]{\small{John Coltrane's Giant Steps head: A testament to harmonic exploration, featuring rapid chord changes in non-diatonic major thirds.}}
    \label{fig:giant_steps}
\end{figure}

>>>>

This thesis posits that enduring, high-level musical elements persist regardless of sonic changes and constantly evolve within their context, mirroring the nature of sheet music's symbolic domain. A single waveform can correspond to a finite set of representations. Still, one music piece can yield infinite interpretations, varying by factors such as instruments and performance styles, to name a very few.

The premise of our work is that regardless of the waveform or production style, abstract high-level features can be discerned, offering objective and thorough insight into a composition's musical content, analogous to analyzing sheet music for its basic behaviors \ref{fig:Wandrers Nachtlied, Op. 4, D. 224}.

%%%%%%%%%%%%%%%%%%

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption[Excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert.]{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. Passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and their tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

%%%%%%%%%%%%%%%%%%

@@

Despite differing cultural contexts, styles, time and key signatures, and sonic properties, we suggest a commonality between figures \ref{fig:mahler} and \ref{fig:giant_steps}. Both pieces display the same 'musicological flavor', derived from their melodic and harmonic contour envelopes through non-diatonic major thirds.

@@

\begin{figure}[ht]
    \centering
    \scalebox{0.9}{
    \includegraphics[width=\textwidth]{figures/images/giant steps score.png}
        }
    \caption[Giant Steps]{\small{John Coltrane's Giant Steps head, featuring rapid chord changes in non-diatonic major thirds.}}
    \label{fig:giant_steps}
\end{figure}

### 2023-06-29 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music \footnote{Throughout this thesis, "music" refers to the Western tradition, defined by the conventions, practices, and aesthetics that primarily evolved in Europe and North America. This term's usage does not negate the diversity and richness of other musical traditions worldwide. Instead, it specifies this study's concentration on a particular cultural context.}, a cornerstone of human culture and history, has developed hand in hand with our societies since ancient times. However, the study of music is complex due to its inherent subjectivity, ever-changing ground truth, and the array of styles and cultural contexts it encompasses.

Prior research emphasizes the necessity of selecting suitable audio features to distinguish and label unique music segments for music segmentation tasks. The challenge lies in acquiring annotated data for these feature transformations, which can be time-consuming and expensive. In response, scholars have turned to unsupervised deep learning using readily available music audio \cite{unsupervisedlearndeepfeat, GrillMUSICANNOTATIONS}. This method has substantially improved the generalization capacity of machine-listening models in downstream applications. Among others, segmentation algorithms have seen significant enhancements and delivered state-of-the-art results \cite{Hernandez-Olivan2021MusicFeatures, Li2023MERT:Training}.

Beyond applications in music structure, the evolution of neural networks has empowered the creation of condensed numerical representations encapsulating crucial musical traits. This evolution has simplified the computational process for various task-specific elements, including but not limited to:

@@

Furthermore, deep audio embeddings offer the benefit of transferability, establishing a foundation for multiple tasks. Compared to training a model from scratch, this approach saves computational resources and time \cite{}.

Given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise for Music Information Retrieval (MIR) downstream tasks, including boundary detection. However, it is yet to be determined whether a general-purpose audio representation can successfully emulate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have demonstrated generalizability across as many as fourteen music-understanding tasks \cite{Li2023MERT:Training}.

>>>>

Music \footnote{Throughout this thesis, "music" refers to the Western tradition, defined by the conventions, practices, and aesthetics that primarily evolved in Europe and North America. This term's usage does not negate the diversity and richness of other musical traditions worldwide. Instead, it specifies this study's concentration on a particular cultural context.}, a cornerstone of human culture and history, has developed hand in hand with our societies since ancient times. However, its study presents unique challenges due to its subjectivity, fluctuating ground truth, and the vast array of styles and cultural contexts it encompasses.

Prior research emphasizes the necessity of selecting suitable audio features to distinguish and label unique music segments for music segmentation tasks. The challenge lies in acquiring annotated data for these feature transformations, which can be time-consuming and expensive. In response, scholars have turned to unsupervised deep learning using readily available music audio \cite{deepfeaturesegment, Grill2015MusicAnnotations}. This method has substantially improved the generalization capacity of machine-listening models in downstream applications. Among others, segmentation algorithms have seen significant enhancements and delivered state-of-the-art results \cite{Hernandez-Olivan2021MusicFeatures, Li2023MERT:Training}.

Beyond applications in music structure analysis, the evolution of neural networks has empowered the creation of condensed numerical representations encapsulating crucial musical traits, technically known as representation learning. This evolution has simplified the computational process for various task-specific elements, including but not limited to:

@@

Furthermore, deep audio embeddings offer the benefit of transferability, establishing a foundation for multiple tasks. Compared to training a model from scratch, this approach saves computational resources and time \cite{transferMIR2013, CifkaDeepTransfer}.

Given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise for MIR downstream tasks, including boundary detection. However, it is yet to be determined whether a general-purpose audio representation can successfully emulate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have demonstrated generalizability across many music-understanding tasks \cite{Li2023MERT:Training}.

Yet, despite these strides, deep learning's role in music understanding remains relatively nascent, with scarce work on deep music representations, a dearth of large-scale datasets, and a lack of a universal community-driven benchmark \cite{Yuan2023MARBLE:Evaluation}.

### 2023-06-29 introduction/Subsections/Motivation.tex
Updates from Overleaf

The collaboration aims to deepen our understanding of music's fundamental structures, which could enhance ES's technical offerings and further advancements in music information retrieval (MIR) research.

As a musician, enthusiast, and scholar, my passion for music has driven me to investigate the foundational elements of musical composition, various creation techniques, and their intricate relationships. I am dedicated to extracting valuable information embedded within music across all domains, with a particular focus on audio and symbolic. Musicians who endeavor to bridge the gap and uncover musical truths within the tonal paradigm, such as Heinrich Schenker \cite{Komar1959SchenkersStructure}, or those who challenge it, like Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell}, and Ernst Levy \cite{LevyAHarmony}, have been a continual source of inspiration. They aim to identify abstract concepts that reinforce or disrupt the tonal foundation, advancing the tonal landscape and providing a solid base for musicians' growth, development, and understanding of tonal and atonal paradigms.

Advancements in AI research, building on the theoretical groundwork laid by pioneers such as Alan Turing and Claude Shannon, have led to significant breakthroughs \cite{Vaswani2017AttentionNeed}. With AI being applied broadly in popular products \cite{OpenAI2023GPT-4Report}, researchers and corporations alike must stay abreast of this rapidly evolving field. The emergence of self-supervised models capable of learning embedding spaces to retrieve musical content from audio signals presents new opportunities. These models, which autonomously extract high-level semantic information from raw audio data, can potentially transform multiple aspects of the music industry. Industrially, these embedding spaces can be used to devise innovative products, enhance user experiences for content creators, and stimulate innovation and collaboration across the industry.

Traditionally, MIR research has focused primarily on technical aspects, perhaps overlooking relevant areas such as perception, cognition, music theory, musicology, aural skills, and high-level perceptual concepts. Purely mathematical approaches may not accurately represent human music perception and could miss subtle variations in timing or harmonic relationships essential to the musical experience. Therefore, MIR research should strive for a balanced approach, recognizing the interdisciplinary nature of music and the importance of areas beyond digital signal processing.

%%%%%%%%%%%%%%%%%%

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption[Excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert.]{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. Passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and their tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

>>>>

The collaboration aims to deepen our understanding of music's fundamental structures, which could enhance ES's technical offerings and further advancements in MIR research.

As a musician, my passion for music has driven me to investigate the foundational elements of musical composition, various creation techniques, and their intricate relationships. I am dedicated to extracting valuable information embedded within music across all domains, focusing on audio and sheet music (symbolic domain). Musicians who endeavor to bridge the gap and uncover musical truths within the tonal paradigm, such as Heinrich Schenker \cite{Komar1959SchenkersStructure}, or those who challenge it, like Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell}, and Ernst Levy \cite{LevyAHarmony}, have been a continual source of inspiration. They aim to identify abstract concepts that reinforce or disrupt the tonal foundation, advancing the tonal landscape and providing a solid base for musicians' growth, development, and understanding of tonal and atonal paradigms.

Advancements in AI research, building on the theoretical groundwork laid by pioneers such as Alan Turing and Claude Shannon, have led to significant breakthroughs \cite{Vaswani2017AttentionNeed}. With AI being applied broadly in skyrocketing popular products \cite{OpenAI2023GPT-4Report}, researchers and corporations must stay abreast of this rapidly evolving field. The emergence of self-supervised models capable of learning embedding spaces to retrieve musical content from audio signals presents new opportunities. These models, which autonomously extract information from audio data, can potentially transform multiple aspects of the music industry. Industrially, these embedding spaces can be used to devise innovative products, enhance user experiences for content creators, and stimulate innovation and collaboration across the industry.

Traditionally, MIR research has focused primarily on technical aspects, perhaps overlooking relevant areas such as perception, cognition, music theory, musicology, aural skills, and high-level perceptual concepts. Purely mathematical approaches may not accurately represent human music perception and could miss subtle details essential to the musical experience \cite{neglecteduser2013}. Therefore, MIR research should strive for a balanced approach, recognizing the interdisciplinary nature of music and the importance of areas beyond digital signal processing.

### 2023-06-29 introduction/Subsections/Objectives.tex
Updates from Overleaf

Can deep audio embeddings be learned from reordering and scrambling sequences of musical information, and can audio effects be added to improve unsupervised music boundary detection?

@@

This study aims to develop and learn a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output - usually a fixed-size vector - produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a piece of music that integrates the high-level relationships between musical elements as they unfold over time.

@@

This approach seeks to replicate human auditory capabilities by understanding and identifying abstract and semantic musical elements independent of their sonic qualities.

>>>>

Can deep audio embeddings be learned from reordering, scrambling, and augmenting via audio effects sequences of musical information to improve unsupervised music boundary detection?

@@

Our research aims to develop and learn a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output - usually a fixed-size vector - produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a piece of music that integrates the high-level relationships between musical elements as they unfold over time.

@@

This approach replicates human auditory capabilities by understanding and identifying abstract and semantic musical elements independent of their sonic qualities.

### 2023-06-29 introduction/Subsections/Previous work.tex
Updates from Overleaf

Related research such as \cite{SalamonDeepSegmentation} and \cite{deepfeaturesegment} explore how the use of audio embeddings learned via unsupervised methods can enhance the performance of music segmentation algorithms leveraging the power of deep learning and data-driven feature learning to improve upon traditional, manually-engineered methods of music segmentation.

>>>>

Related research such as \cite{deepfeaturesegment, SalamonDeepSegmentation} explore how the use of audio embeddings learned via unsupervised methods can enhance the performance of music segmentation algorithms leveraging the power of deep learning and data-driven feature learning to improve upon traditional, manually-engineered methods of music segmentation.

### 2023-06-29 results/subsections/tables.tex
Updates from Overleaf

The evaluation results have been obtained using the standard and most commonly used MIR's tools and frameworks such as MSAF \cite{NietoMSAF:FRAMEWORK} for the segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} as evaluation ground truth and $mir_eval$ \cite{RaffelMir_eval:METRICS} for metric computation.

@@

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
\hline
\textbf{Technique} & \textbf{Training Dataset} & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\hline
MFCC & -- & 0.173 & 0.185 & 0.172 \\
\hline
Embeddiogram & GTZAN Dataset & 0.228 & 0.171 & 0.185 \\
\hline
Embeddiogram & MSD Dataset & \textbf{0.335} & 0.276 & 0.288 \\
\hline
PCP & -- & 0.311 & 0.324 & 0.305 \\
\hline
Tonnetz & -- & 0.312 & 0.312 & 0.300 \\
\hline
CQT & -- & 0.311 & \textbf{0.339} & \textbf{0.312} \\
\hline
\end{tabularx}
\caption[Comparison of precision, recall, and f-measure for different audio features]{Comparison of precision, recall, and f-measure for different features using the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}
\label{tab:comparison}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & \textbf{0.523} \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. Original source \cite{Hernandez-Olivan2021MusicFeatures}}}
\label{tab:comparison_table}
\end{table}

>>>>

The evaluation results have been obtained using the standard and most commonly used MIR's tools and frameworks, such as default values of MSAF \cite{MSAF} for the segmentation algorithms \cite{sf}, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} as evaluation ground truth and $mir_eval$ \cite{RaffelMir_eval:METRICS} for metric computation.

@@

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
\hline
\textbf{Technique} & \textbf{Training Dataset} & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\hline
MFCC & -- & 0.173 & 0.185 & 0.172 \\
\hline
Embeddiogram & GTZAN Dataset & 0.228 & 0.171 & 0.185 \\
\hline
Embeddiogram & MSD Dataset & \textbf{0.335} & 0.276 & 0.288 \\
\hline
PCP & -- & 0.311 & 0.324 & 0.305 \\
\hline
Tonnetz & -- & 0.312 & 0.312 & 0.300 \\
\hline
CQT & -- & 0.311 & \textbf{0.339} & \textbf{0.312} \\
\hline
\end{tabularx}
\caption[Metric comparison for different audio features]{Comparison of precision, recall, and f-measure for different features obtained using the SF algorithm \cite{sf} on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}
\label{tab:comparison}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
\cite{deepfeaturesegment}, 2019 & XXXX & CNN  & \textbf{0.535} ± 0.15 \\
\addlinespace
\cite{sf}, 2021 & XXXX & CNN  & XXX \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}}}
\label{tab:comparison_table}
\end{table}
